rbcL community comparison

### 2019-11-04 code/rbcl_community_compare.Rmd
minor

# run NMDS ordination on results 
nmds <- metaMDS(unifrac, autotransform = F)
```

@@

```{r setup_rel_abund_barplot, include=F}
plot_data <- rel_abund %>% 
  left_join(otus, by="asv_id") %>% 
  group_by(sample_id, otu_name) %>% 
  summarise(rel_abund = sum(rel_abund)) %>%   
  left_join(field, by="sample_id") %>%
  ungroup() %>% 
  mutate(rel_abund = if_else(rel_abund <0.01, 0.01, rel_abund)) # artificially boost rel abund <1% to 1% so its visible on plot

# check rel abund sums ~= 1 (will be slightly higher, due to rounding)
plot_data %>%
  group_by(alias) %>%
  summarise(sum(rel_abund))
```

@@

otu_names %>% distinct(otu_name)

@@

barplot <- ggplot(plot_data, aes(x=alias %>% fct_reorder(elev_m), y=rel_abund)) +
  geom_bar(aes(fill=otu_name %>% fct_relevel(otu_ord)), stat="identity") + # classic stacked relative abundance plot
  coord_flip() +
  facet_grid(~otu_name %>% fct_relevel(otu_ord), switch = "x") +
  labs( x="Sample ID", tag="B") +
  cust_theme
barplot

@@

```{r faiths, echo=F}
re_wide_mat <- ra_wide %>% as.matrix()

faiths <- re_wide_mat %>% 
  picante::pd(rbcl.tree) %>% 
  # attach elevation data
  rownames_to_column("sample_id") %>% 
  left_join(field, by="sample_id")

@@

```{r}
faith_elev_plot <- ggplot(faiths, aes(y=PD, x=elev_m))+
  geom_point() +
  geom_smooth(method = "lm")+
  geom_text_repel(aes(label=alias))+
  labs(title="Phylogenetic diversity of snow algae rbcL vs elevation",y="Faith's phylogenetic diversity (PD)",x="Elevation (m)")
faith_elev_plot
```

@@

```{r output, include=F}
ggsave(filename = here("manuscript/figs/3_community/combined_fig_3.pdf"), plot=combined_fig_3, height=10, width=7)
ggsave(filename = here("manuscript/figs/supplementary/diversity.png"), plot = faith_date_plot, height = 5, width = 7)
```

>>>>

# run NMDS ordination on results 
set.seed(123)
nmds <- metaMDS(unifrac, autotransform = F)
```

@@

```{r setup_rel_abund_barplot, include=F}
otu_rel_abund <- rel_abund %>% 
  left_join(otus, by="asv_id") %>% 
  group_by(sample_id, otu_name) %>% 
  summarise(rel_abund = sum(rel_abund)) %>%   
  left_join(field, by="sample_id") %>%
  ungroup()

otu_rel_abund_round_to_1_percent <- otu_rel_abund %>% 
  mutate(rel_abund = if_else(rel_abund <0.01, 0.01, rel_abund)) # artificially boost rel abund <1% to 1% so its visible on plot

# check rel abund sums ~= 1 (will be slightly higher, due to rounding)
otu_rel_abund_round_to_1_percent %>%
  group_by(alias) %>%
  summarise(sum(rel_abund))
```

@@

otus %>% distinct(otu_name)

@@

barplot <- ggplot(otu_rel_abund_round_to_1_percent, aes(x=alias %>% fct_reorder(elev_m), y=rel_abund)) +
  geom_bar(aes(fill=otu_name %>% fct_relevel(otu_ord)), stat="identity") + # classic stacked relative abundance plot
  coord_flip() +
  facet_grid(~otu_name %>% fct_relevel(otu_ord), switch = "x") +
  labs( x="Sample ID", tag="B") +
  cust_theme
barplot

@@

Faith's Phylogenetic Diversity, sum of branch lengths in a sample's phylogenetic tree (does not take relative abundance into account)
```{r faiths, echo=F}
ra_wide_mat <- ra_wide %>% as.matrix()

faiths <- ra_wide_mat %>% 
  picante::pd(rbcl.tree) %>% 
  # attach elevation data
  rownames_to_column("sample_id") %>% 
  left_join(field, by="sample_id")

@@

```{r, echo=F}
faith_elev_plot <- ggplot(faiths, aes(y=PD, x=elev_m))+
  geom_point() +
  geom_smooth(method = "lm")+
  geom_text_repel(aes(label=alias))+
  labs(title="Phylogenetic diversity of snow algae rbcL vs elevation",y="Faith's phylogenetic diversity (PD)",x="Elevation (m)")
faith_elev_plot
```

Shannon and Simpson diversity indices (these take relative abundance into account, but require OTUs)
```{r, include=F}
otu_rel_abund_mat <- otu_rel_abund %>% 
  select(sample_id, otu_name, rel_abund) %>% 
  pivot_wider(names_from = otu_name, values_from = rel_abund, values_fill = list(rel_abund=0)) %>% 
  column_to_rownames("sample_id") %>% 
  as.matrix()

# use ra_wide_mat for asv based diversity
shannon <- otu_rel_abund_mat %>% 
  diversity(index = "shannon") %>% 
  enframe() %>% 
  dplyr::rename(sample_id = name, shannon = value)
simpson <- otu_rel_abund_mat %>% 
  diversity(index = "simpson") %>% 
  enframe() %>% 
  dplyr::rename(sample_id = name, simpson = value)
otu_diversity <- shannon %>% 
  left_join(simpson, by="sample_id") %>% 
  left_join(field, by="sample_id")

# plot diversity of samples
ggplot(otu_diversity, aes(x=elev_m, y=shannon))+
  geom_point() +
  geom_text_repel(aes(label=alias)) +
  geom_smooth(method = "lm")
```

Chao
```{r}
# otu_rel_abund_mat %>% 
#   mutate
#   # head(1) %>% 
#   as.matrix() %>% 
#   fossil::chao1()
```

@@

```{r output, include=F}
ggsave(filename = here("manuscript/figs/3_community/combined_fig_3.png"), plot=combined_fig_3, height=10, width=7)
ggsave(filename = here("manuscript/figs/supplementary/diversity.png"), plot = faith_date_plot, height = 5, width = 7)
```
